---
title: "A biased comparsion of JSON packages in R"
output: html_document
---

```{r echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(comment="")
options(width = 120)
#make sure we compare most recent versions
stopifnot(packageVersion("rjson") >= "0.2.14")
stopifnot(packageVersion("RJSONIO") >= "1.3-0")
stopifnot(packageVersion("jsonlite") >= "0.9.12")
```

Some simple examples to compare behavior and performance of JSON packages in R.

### Data frame

```{r}
all.equal(mtcars, rjson::fromJSON(rjson::toJSON(mtcars)))
all.equal(mtcars, RJSONIO::fromJSON(RJSONIO::toJSON(mtcars)))
all.equal(mtcars, jsonlite::fromJSON(jsonlite::toJSON(mtcars)))
```

### Matrix

```{r}
(x <- matrix(1:6, 2))
rjson::fromJSON(rjson::toJSON(x))
RJSONIO::fromJSON(RJSONIO::toJSON(x))
jsonlite::fromJSON(jsonlite::toJSON(x))
```

### Lists

```{r}
x <- list(foo = 123, bar= 456)
rjson::fromJSON(rjson::toJSON(x))
RJSONIO::fromJSON(RJSONIO::toJSON(x))
jsonlite::fromJSON(jsonlite::toJSON(x))
```

### Missing values

```{r}
rjson::fromJSON(rjson::toJSON(c(1,2,NA,4)))
RJSONIO::fromJSON(RJSONIO::toJSON(c(1,2,NA,4)))
jsonlite::fromJSON(jsonlite::toJSON(c(1,2,NA,4)))
```

### Escaping

```{r}
x <- list("\b\f\n\r\t" = "\b\f\n\r\t")
identical(x, rjson::fromJSON(rjson::toJSON(x)))
identical(x, RJSONIO::fromJSON(RJSONIO::toJSON(x)))
identical(x, jsonlite::fromJSON(jsonlite::toJSON(x)))
```

### Parser error handling

```{r, error=TRUE}
rjson::fromJSON('[1,2,boo",4]')
RJSONIO::fromJSON('[1,2,boo",4]')
jsonlite::fromJSON('[1,2,boo",4]')
```

### Unicode

```{r, error=TRUE}
json = '["\\u5bff\u53f8","Z\\u00fcrich", "\\u586B"]'
rjson::fromJSON(json)
RJSONIO::fromJSON(json)
jsonlite::fromJSON(json)
```

### Prettify, validate

Only RJSONIO and jsonlite have functionality to validate or prettify JSON:

```{r}
x <- list(foo = c("hi", "hello"), bar=1:3)
cat(RJSONIO::toJSON(x, pretty = TRUE))
cat(jsonlite::toJSON(x, pretty = TRUE))
RJSONIO::isValidJSON(RJSONIO::toJSON(x), asText = TRUE)
jsonlite::validate(jsonlite::toJSON(x))
```

### Digits

RJSONIO uses significant digits, whereas jsonlite uses decimal digits. rjson does not support this.

```{r}
rjson::toJSON(pi)
RJSONIO::toJSON(pi, digits=4)
jsonlite::toJSON(pi, digits=4)
```


### Controlling simplification

Only RJSONIO and jsonlite give control over vector simplification. rjson always simplifies if possible.

```{r}
json <- '[1,2,3,4]'
RJSONIO::fromJSON(json)
RJSONIO::fromJSON(json, simplify = FALSE)
jsonlite::fromJSON(json)
jsonlite::fromJSON(json, simplifyVector = FALSE)
rjson::fromJSON(json)
```

### Performance 1: generating JSON

Comparing performance is a bit difficult because different packages do different things. The `rjson` package has no options to control conversion, so the only way to benchmark common functionality is by trying to mimic `rjson`:

```{r}
library(microbenchmark)
data(diamonds, package="ggplot2")
microbenchmark(
  rjson::toJSON(diamonds),
  RJSONIO::toJSON(diamonds),
  jsonlite::toJSON(diamonds, dataframe = "column"),
  times = 10
)
```

What if we throw some missing values in the mix:

```{r}
diamonds2 <- diamonds
diamonds2[1,] <- NA;
microbenchmark(
  rjson::toJSON(diamonds2),
  RJSONIO::toJSON(diamonds2),
  jsonlite::toJSON(diamonds2, dataframe = "column"),
  times = 10
)
```

### Performance 2: Parsing JSON

For parsing, comparing performance gets even trickier because rjson does not give any control over simplification. The following settings in RJSONIO and jsonlite are roughly equivalent to rjson:

```{r}
json <- rjson::toJSON(diamonds)
microbenchmark(
  rjson::fromJSON(json),
  RJSONIO::fromJSON(json, simplifyWithNames=FALSE),
  jsonlite::fromJSON(json, simplifyDataFrame = FALSE, simplifyMatrix = FALSE),
  times = 10
)
```


